Global Assessments of Fishery Status need Better Data more than Better Models

### 2020-10-02 documents/ovando-etal-assessing-global-fisheries.Rmd
merging back to master

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random quess. Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

@@

We should of course prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database. This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone. However, our results demonstrate that we must prioritize aggregation of long-running but less publicly-accessible survey data, or expect to wait for some time until new surveys are capable of yielding meaningful information on their own about stock status. Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//). Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.

For many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible in the near future. Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies. While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health. While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results at the macro scale, we attribute this largely to two difficulties.

>>>>

---
title: "Global Assessments of Fishery Status need Better Data more than Better Models"
author: 
  - Daniel Ovando
  - Ray Hilborn
  - Cole Monnahan
  - Merrill Rudd
  - Rishi Sharma
  - James Thorson
  - Yannick Rousseau
  - Yimin Ye
date: "`r Sys.Date()`"
bibliography: ["../references.bib"]
csl: nature.csl
output: 
  bookdown::word_document2:
    reference_docx: word-template.docx
  bookdown::pdf_document2: default
params:
  results_name: ["v0.5"]
  min_years_catch: [25]
abstract: |
  Assessments of the global state of fisheries play an important role in shaping the public narrative around ocean health, motivating future directions of research and funding, formulating evidence-based policy and tracking the implementation of the United Nations Sustainable Development Goals. While we have reliable estimates of stock status for fisheries accounting for 50% of global catch, our knowledge of the state of the remaining 50%, the worlds 'unassessed' fisheries, is poor. Numerous high-profile publications featuring a range of statistical methods have produced estimates of the global status of these unassessed fisheries, but limited quantity and quality of data along with methodological differences have produced counterintuitive and conflicting results. This is especially true in areas such as Southeast Asia and Africa which are also regions that have high dependence on fishery resources. How can we effectively estimate the state of global fishery sustainability and track progress towards the Sustainable Development Goals targets? We developed a flexible assessment model to assess the value of different kinds, quantities, and quality of data in improving estimates of fishery stock status. We then explore avenues for obtaining potentially impactful data, including through the use of local expert opinion through Fisheries Management Index scores, and increasingly available but historically underutilized data such as trawl footprints and effort data. These data are then used to illustrate how different types of information paint starkly different pictures of the state of fisheries around the world, and to identify priority data types for future collection. Our results provide further evidence that relying on catch data alone results in inaccurate classification of fishery status, often performing only as well or worse than a random guess. Obtaining accurate estimates of stock status for the world's unassessed fisheries depends on prioritizing the collection of high-priority data at a global scale, not on the development of new modeling methods alone.
always_allow_html: true
linkcolor: blue
---

@@

# check in call notes

Ray: setting the stage: what we want is a way to assess the status of stocks and see how we're doing relative to SDG.

We need to discuss a sampling strategy: how can we get down to a set of priority stock. A step of expert knowledge to refine what stocks you would focus in on.

Regressions based approach, focusing in effort on stocks

Basically a paragraph laying out that we need to do this, and discussing how we might go about doing that. Equal weight to each stock vs proportional to biomass

add in a bit more to the discussion on guidelines

add in error in the index for the VOI analysis

Apples and apples comparisons between model and design developments. How much is the value of the effort data relative to actually improving the population model. Production model is ill-suited to some of these questions.

fishing effort uncertainty of the effort

how credible is the information

probably don't need to bother with the SAUP

stock assessments

the stock will collapse next year

nansen data

@@

We should of course prioritize collection of survey data worldwide, high-quality synthesis and standardization of these data, as well as their collection in an international database. This assessment assumes that we have perfect knowledge of for example F/F~MSY~, when in fact any estimates derived from for example length-based methods are likely to be highly error-prone. However, our results demonstrate that we must prioritize aggregation of long-running but less publicly-accessible survey data, or expect to wait for some time until new surveys are capable of yielding meaningful information on their own about stock status. Repositories of fishery-independent survey data would be immensely beneficial, such as those maintained by [FishStat](https://james-thorson.github.io//). Recent research confirms that there are bottom trawl data to support analysis of biomass-trends since 2001 and potentially earlier in many regions [@maureaud2020], and survey data are available for more stocks than have previously had stock assessments.  xx bring this home xx

For many areas of the world effective fishery independent estimates of stock abundance are unlikely to be feasible in the near future. XXbe more specific here, and clearXX. .Effort databases such as @rousseau2019, can be combined with catch histories to to construct fishery-dependent abundance indicies. While there are substantial difficulties and biases that can arise from improper use of fishery-dependent abundance data, improved understanding of the evolution of global fishing effort could be very useful in improving global estimates of fishery health. While we did include effort reconstructions in our analysis, and did not find it to provide substantial improvements in assessment results at the macro scale, we attribute this largely to two difficulties.
